Informing policy via dynamic models: Eliminating cholera in Haiti

### 2022-08-05 ms.Rnw
Refactored Model 1 scenario code

RUN_LEVEL <- 1

@@

myround<- function (x, digits = 1) {
  # taken from the broman package
  if (digits < 1)
    stop("This is intended for the case digits >= 1.")
  if (length(digits) > 1) {
    digits <- digits[1]
    warning("Using only digits[1]")
  }
  tmp <- sprintf(paste("%.", digits, "f", sep = ""), x)
  zero <- paste0("0.", paste(rep("0", digits), collapse = ""))
  tmp[tmp == paste0("-", zero)] <- zero
  tmp
}

@@

dateToYears <- function(date, origin = as.Date("2014-01-01"), yr_offset = 2014) {
  # This function converts a date to a decimal representation
  #
  # ex: "1976-03-01" -> 1976.163

@@

stew(
  file = "model2/output/model2_fit.rda", {h2_fit <- fit_haiti2()}
)

# load('model2/output/model2_fit_params.rda')
# p2_epi <- h2_epi_params
# h2_fit$end_params <- h2_end_params
# rm(h2_epi_params, h2_end_params)
# gc()

@@

<<Load Model 1 VaccScen Sims, message=FALSE, include=FALSE>>=
load('model1/vacscens/sims_S0.rda')
mod1_V0_sims <- agg_mod1_sims(sims)
mod1_V0_probs <- get_elimProbs(sims, model = 1)
rm(sims)
gc()

load('model1/vacscens/sims_S2.rda')
mod1_V1_sims <- agg_mod1_sims(sims)
mod1_V1_probs <- get_elimProbs(sims, model = 1)
rm(sims)
gc()

load('model1/vacscens/sims_S4.rda')
mod1_V2_sims <- agg_mod1_sims(sims)
mod1_V2_probs <- get_elimProbs(sims, model = 1)
rm(sims)
gc()

load('model1/vacscens/sims_S3.rda')
mod1_V3_sims <- agg_mod1_sims(sims)
mod1_V3_probs <- get_elimProbs(sims, model = 1)
rm(sims)
gc()

load('model1/vacscens/sims_S1.rda')
mod1_V4_sims <- agg_mod1_sims(sims)
mod1_V4_probs <- get_elimProbs(sims, model = 1)
rm(sims)
gc()
@

@@

stew(
  file = "model2/output/VaccinationScenarios.rda", {
    mod2_VaccScenarios <- haiti2_vaccScenario(
      end_params = h2_fit$end_params,
      epi_params = h2_fit$epi_params
    )
  }
)
@

>>>>

RUN_LEVEL <- 1
rl_dir <- paste0("run_level_", RUN_LEVEL, "/")

for (i in 1:3) {
  if (!dir.exists(paste0("model", i, '/', rl_dir))) {
    dir.create(paste0("model", i, '/', rl_dir), recursive = TRUE)
  }
}

@@

myround <- function(x, digits = 1) {
  # taken from the broman package
  if (digits < 1)
    stop("This is intended for the case digits >= 1.")
  if (length(digits) > 1) {
    digits <- digits[1]
    warning("Using only digits[1]")
  }
  tmp <- sprintf(paste("%.", digits, "f", sep = ""), x)
  zero <- paste0("0.", paste(rep("0", digits), collapse = ""))
  tmp[tmp == paste0("-", zero)] <- zero
  tmp
}

@@

# TODO: Remove these functions and replace with lubridate functions (also in the package).
dateToYears <- function(date, origin = as.Date("2014-01-01"), yr_offset = 2014) {
  # This function converts a date to a decimal representation
  #
  # ex: "1976-03-01" -> 1976.163

@@

# Fit model 2. Note that this takes the same ammount of time at all run
# levels.
h2_fit <- bake(
  file = paste0("model2/", rl_dir, "model2_fit.rds"), {
    fit_haiti2()
  },
  timing = FALSE
)

@@

<<Load Model 1 VaccScen Sims, message=FALSE, include=FALSE>>=

# Number of particles and simulations to use for pfilter and project from filter, respectively.
h1_scen_NP   <- switch(RUN_LEVEL, 50, 500, 5000)
h1_scen_sims <- switch(RUN_LEVEL, 20, 100,  500)

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
h1_S0_sims <- bake(
  file = paste0("model1/", rl_dir, "S0_sims.rds"), {

    # Create the model
    s0 <- haiti1_joint(vacscen = 'id0')

    # Load coefficients
    coef(s0) <- p1

    # Get samples from filtering distribution
    s0_pf <- pfilter(s0, save.state = TRUE, Np = h1_scen_NP)

    # project vacc scenario, with filtering dist, and model.
    project_from_filter2(mod = s0, PF = s0_pf, nsims = h1_scen_sims)
  },
  timing = FALSE
)

# Clean-up simulations for analysis/visualization
mod1_V0_sims <- agg_mod1_sims(h1_S0_sims)
mod1_V0_probs <- get_elimProbs(h1_S0_sims, model = 1)
rm(h1_S0_sims)

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
h1_S2_sims <- bake(
  file = paste0('model1/', rl_dir, 'S2_sims.rds'), {

    # Create the model
    s2 <- haiti1_joint(vacscen = "id2")

    # Set model parameters that need to be adjusted
    depts <- 2
    h1_par_names <- names(p1)
    h1_params_temp <- c(p1, rep(0.0, 5 * depts))
    names(h1_params_temp) <- c(
      h1_par_names,
      paste0("S", 1:depts, "_0"),
      paste0("E", 1:depts, "_0"),
      paste0("I", 1:depts, "_0"),
      paste0("A", 1:depts, "_0"),
      paste0("R", 1:depts, "_0")
    )
    coef(s2) <- h1_params_temp  # Set model parameters

    # Get samples from filtering distribution for Scenario 2 model:
    s2_pf <- pfilter(s2, save.state = TRUE, Np = h1_scen_NP)

    # project vacc scenario, with filtering dist, and model.
    project_from_filter2(mod = s2, PF = s2_pf, nsims = h1_scen_NP)
  },
  timing = FALSE
)

# Clean-up simulations for analysis/visualization
mod1_V1_sims <- agg_mod1_sims(h1_S2_sims)
mod1_V1_probs <- get_elimProbs(h1_S2_sims, model = 1)
rm(h1_S2_sims)

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
h1_S4_sims <- bake(
  file = paste0('model1/', rl_dir, 'S4_sims.rds'), {

    # Create the model
    s4 <- haiti1_joint(vacscen = "id4")

    # Set model parameters that need to be adjusted
    depts <- 3
    h1_par_names <- names(p1)
    h1_params_temp <- c(p1, rep(0.0, 5 * depts))
    names(h1_params_temp) <- c(
      h1_par_names,
      paste0("S", 1:depts, "_0"),
      paste0("E", 1:depts, "_0"),
      paste0("I", 1:depts, "_0"),
      paste0("A", 1:depts, "_0"),
      paste0("R", 1:depts, "_0")
    )
    coef(s4) <- h1_params_temp  # Set model parameters

    # Get samples from filtering distribution for Scenario 2 model:
    s4_pf <- pfilter(s4, save.state = TRUE, Np = h1_scen_NP)

    # project vacc scenario, with filtering dist, and model.
    project_from_filter2(mod = s4, PF = s4_pf, nsims = h1_scen_NP)
  },
  timing = FALSE
)

# Clean-up simulations for analysis/visualization
mod1_V2_sims <- agg_mod1_sims(h1_S4_sims)
mod1_V2_probs <- get_elimProbs(h1_S4_sims, model = 1)
rm(h1_S4_sims)

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
h1_S3_sims <- bake(
  file = paste0('model1/', rl_dir, 'S3_sims.rds'), {

    # Create the model
    s3 <- haiti1_joint(vacscen = "id3")

    # Set model parameters that need to be adjusted
    depts <- 10
    h1_par_names <- names(p1)
    h1_params_temp <- c(p1, rep(0.0, 5 * depts))
    names(h1_params_temp) <- c(
      h1_par_names,
      paste0("S", 1:depts, "_0"),
      paste0("E", 1:depts, "_0"),
      paste0("I", 1:depts, "_0"),
      paste0("A", 1:depts, "_0"),
      paste0("R", 1:depts, "_0")
    )
    coef(s3) <- h1_params_temp  # Set model parameters

    # Get samples from filtering distribution for Scenario 2 model:
    s3_pf <- pfilter(s3, save.state = TRUE, Np = h1_scen_NP)

    # project vacc scenario, with filtering dist, and model.
    project_from_filter2(mod = s3, PF = s3_pf, nsims = h1_scen_NP)
  },
  timing = FALSE
)

# Clean-up simulations for analysis/visualization
mod1_V3_sims <- agg_mod1_sims(h1_S3_sims)
mod1_V3_probs <- get_elimProbs(h1_S3_sims, model = 1)
rm(h1_S3_sims)

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
h1_S1_sims <- bake(
  file = paste0('model1/', rl_dir, 'S1_sims.rds'), {

    # Create the model
    s1 <- haiti1_joint(vacscen = "id1")

    # Set model parameters that need to be adjusted
    depts <- 10
    h1_par_names <- names(p1)
    h1_params_temp <- c(p1, rep(0.0, 5 * depts))
    names(h1_params_temp) <- c(
      h1_par_names,
      paste0("S", 1:depts, "_0"),
      paste0("E", 1:depts, "_0"),
      paste0("I", 1:depts, "_0"),
      paste0("A", 1:depts, "_0"),
      paste0("R", 1:depts, "_0")
    )
    coef(s1) <- h1_params_temp  # Set model parameters

    # Get samples from filtering distribution for Scenario 2 model:
    s1_pf <- pfilter(s1, save.state = TRUE, Np = h1_scen_NP)

    # project vacc scenario, with filtering dist, and model.
    project_from_filter2(mod = s1, PF = s1_pf, nsims = h1_scen_NP)
  },
  timing = FALSE
)

# Clean-up simulations for analysis/visualization
mod1_V4_sims <- agg_mod1_sims(h1_S1_sims)
mod1_V4_probs <- get_elimProbs(h1_S1_sims, model = 1)
rm(h1_S1_sims)

gc()
@

@@

# Run Vaccination scenarios for Model 2. Note that the code runs the same
# speed for each run-level.
mod2_VaccScenarios <- bake(
  file = paste0("model2/", rl_dir, "VaccinationScenarios.rds"), {
    haiti2_vaccScenario(
      end_params = h2_fit$end_params,
      epi_params = h2_fit$epi_params
    )
  },
  timing = FALSE
)
